Revisiting the Relationship between Short-Chain Fatty Acids, the Microbiota, and Colorectal Tumors

### 2018-08-06 submission/manuscript.Rmd
update with ada suggestions paragraph 4

@@

**SCFA concentrations do not replace important *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* OTUs in Random Forest models built to classify tumors.** SCFA concentrations could improve prediction of tumors based on specific bacterial community structures. Our OTU data can be used in combination with SCFAs to assess whether there is a community dependent context to SCFA classification of tumors. Using the Random Forest algorithm we built models with OTU abundance data or OTU abundances and SCFA concentrations to classify normal versus adenoma and normal versus carcinoma fecal samples. For adenoma and carcinoma models, there was no difference between the median AUC of models with or without SCFA concentrations (P-value `r ifelse(min(pull(disease_rf, bh)) > 0.05, invisible(">"), invisible("<"))` 0.05) [Figure 4A & 4D]. There was little difference between the top 10 most important OTUs, as measured by mean decrease in accuracy, in models using SCFA and OTUs versus OTUs only [Figure 4B-C & 4E-F]. The main reason for this was because SCFA concentrations replaced the information gained by specific OTUs rather than add new information to the model [Figure 4B-C]. The SCFA concentrations also only replaced select OTUs with taxonomic classification to *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* with many OTUs within these taxa remaining in the top 10 [Figure 4B & 4C]. In combination with the previous results on taxa correlations, these observations provide additional evidence that the resident taxa within models that classify tumor are not ones associated with acetate, butyrate, or propionate production.

>>>>

# Load in sig associations and top 10 RF model presence
sig_assoc_top10_model <- read.csv("../data/process/tables/sig_association_and_RF_top10.csv")

@@

**SCFA concentrations do not replace important *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* OTUs in Random Forest models built to classify tumors.** Despite the lack of difference in positive correlations between OTUs and SCFA concentrations between individuals with and without tumors, OTUs associated with SCFA concentrations could still be the most important variables to Random Forest models built to classify patients with adenomas or carcinomas. We tested this by using the Random Forest algorithm to build models with OTU abundance data or OTU abundances and SCFA concentrations to classify normal versus adenoma and normal versus carcinoma fecal samples. With these models we compared whether OTUs with taxonomic classification to *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* remained in the OTU and SCFA model. Additionally, we also compared whether any of the important OTUs that remained also had significant correlations with SCFA concentrations. Both our adenoma and carcinoma models classified patients with a similar degree of success, as measured by the area under the curve (AUC) (P-value `r ifelse(min(pull(disease_rf, bh)) > 0.05, invisible(">"), invisible("<"))` 0.05) [Figure 4A & 4D]. After the addition of SCFA concentrations to the adenoma or carcinoma models, many OTUs with taxonomic classification to *Clostridiales*, *Lachnospiraceae*, and *Ruminococcaceae* remained as important variables to the model [Figure Figure 4B-C & 4E-F]. After adding SCFA concentrations to the adenoma model, there were only `r length(unique(sig_assoc_top10_model %>% filter(model == "adn_full") %>% pull(otu)))` OTUs significantly associated with SCFA concentration that were part of the top 10 most important variables and both were postively associated with acetate, butyrate, or propionate concentrations. When SCFA concentrations were added to the carcinoma model, only `r length(unique(sig_assoc_top10_model %>% filter(model == "crc_full") %>% pull(otu)))` OTU that was associated with SCFA concentration remained as part of the top 10 most important variables and it was negatively associated with acetate and butyrate concentrations. In combination with the previous results on OTU correlations, these observations provide additional evidence that the resident taxa that are associated with protection against tumorigeneis are not ones associated with acetate, butyrate, or propionate production.

### 2018-08-06 submission/manuscript.tex
update with ada suggestions paragraph 4

\textbf{SCFA concentrations do not replace important
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
OTUs in Random Forest models built to classify tumors.} SCFA
concentrations could improve prediction of tumors based on specific
bacterial community structures. Our OTU data can be used in combination
with SCFAs to assess whether there is a community dependent context to
SCFA classification of tumors. Using the Random Forest algorithm we
built models with OTU abundance data or OTU abundances and SCFA
concentrations to classify normal versus adenoma and normal versus
carcinoma fecal samples. For adenoma and carcinoma models, there was no
difference between the median AUC of models with or without SCFA
concentrations (P-value \textgreater{} 0.05) {[}Figure 4A \& 4D{]}.
There was little difference between the top 10 most important OTUs, as
measured by mean decrease in accuracy, in models using SCFA and OTUs
versus OTUs only {[}Figure 4B-C \& 4E-F{]}. The main reason for this was
because SCFA concentrations replaced the information gained by specific
OTUs rather than add new information to the model {[}Figure 4B-C{]}. The
SCFA concentrations also only replaced select OTUs with taxonomic
classification to \emph{Clostridiales}, \emph{Lachnospiraceae}, and
\emph{Ruminococcaceae} with many OTUs within these taxa remaining in the
top 10 {[}Figure 4B \& 4C{]}. In combination with the previous results
on taxa correlations, these observations provide additional evidence
that the resident taxa within models that classify tumor are not ones
associated with acetate, butyrate, or propionate production.

>>>>

\textbf{SCFA concentrations do not replace important
\emph{Clostridiales}, \emph{Lachnospiraceae}, and \emph{Ruminococcaceae}
OTUs in Random Forest models built to classify tumors.} Despite the lack
of difference in positive correlations between OTUs and SCFA
concentrations between individuals with and without tumors, OTUs
associated with SCFA concentrations could still be the most important
variables to Random Forest models built to classify patients with
adenomas or carcinomas. We tested this by using the Random Forest
algorithm to build models with OTU abundance data or OTU abundances and
SCFA concentrations to classify normal versus adenoma and normal versus
carcinoma fecal samples. With these models we compared whether OTUs with
taxonomic classification to \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} remained in the OTU
and SCFA model. Additionally, we also compared whether any of the
important OTUs that remained also had significant correlations with SCFA
concentrations. Both our adenoma and carcinoma models classified
patients with a similar degree of success, as measured by the area under
the curve (AUC) (P-value \textgreater{} 0.05) {[}Figure 4A \& 4D{]}.
After the addition of SCFA concentrations to the adenoma or carcinoma
models, many OTUs with taxonomic classification to \emph{Clostridiales},
\emph{Lachnospiraceae}, and \emph{Ruminococcaceae} remained as important
variables to the model {[}Figure Figure 4B-C \& 4E-F{]}. After adding
SCFA concentrations to the adenoma model, there were only 2 OTUs
significantly associated with SCFA concentration that were part of the
top 10 most important variables and both were postively associated with
acetate, butyrate, or propionate concentrations. When SCFA
concentrations were added to the carcinoma model, only 1 OTU that was
associated with SCFA concentration remained as part of the top 10 most
important variables and it was negatively associated with acetate and
butyrate concentrations. In combination with the previous results on OTU
correlations, these observations provide additional evidence that the
resident taxa that are associated with protection against tumorigeneis
are not ones associated with acetate, butyrate, or propionate
production.
